(tentative) Rotational Spectral Unmixing of Exoplanets

### 2016-10-17 paper/main.tex
figure size reduced
text reorganized, figure added

\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{rotating}
\usepackage{natbib}
\usepackage{enumerate}
\usepackage{multirow}
\usepackage{array}
\usepackage{appendix}
\usepackage{comment}
\usepackage{color,xcolor}
\usepackage{url}
\usepackage{hyperref}
\hypersetup{colorlinks,linkcolor={blue!50!black},citecolor={blue!50!black},urlcolor={blue!50!black}}
\allowdisplaybreaks[1]
\bibliographystyle{apj}
\renewcommand{\bibname}{References}

@@

@@

By definition, the area fraction $f$ and reflectance spectra $s$ are subject to the following conditions:
%%%
\begin{eqnarray}
\begin{cases}
\;\; 0 < s_{kj} < 1 \;\;\; & \mbox{for any $k$, $j$} \\
\;\; 0 < \fast_{lk} < 1 \;\;\; & \mbox{for any $l$, $k$} \label{eq:cond_f_ast}\\
\;\; \sum_k \fast_{lk} = 1 & \mbox{for any $l$} 
\end{cases}
\end{eqnarray}
%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[bh]
\caption{Indexes}
\begin{center}
\begin{tabular}{ccc} \hline \hline
Name & Symbol & Maximum \\ \hline
Observation Time & $i$ & I \\
Band & $j$ & J  \\
Surface Type & $k$ & K  \\
Longitudinal Slice  & $l$ & L \\ \hline
\end{tabular}
\end{center}
\label{tab:index}
\end{table}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The time variability of the apparent covering fraction $\fast $ due to spin rotation is related to the surface inhomogeneity along the equator. Approximately, $\fast _{ik}$ may be written as the weighted summation of the area fraction of $k$-th surface type in each of (the finite number $L$ of) longitudinal slices, i.e.,
%%%
\begin{eqnarray}
\fast _{ik} &=& \frac{ \int_{{\rm IV}_{i}} f_k (\vec \Omega ) \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }{ \int_{{\rm IV}_i}  \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }  \\
&=& \sum_l \frac{ \int_{{\rm IV}_{il}}  f_k (\vec \Omega ) W (\vec \Omega  ) \cdot d\vec \Omega }{ \int_{{\rm IV}}  W (\vec \Omega ) \cdot d\vec \Omega } \\
&\sim & \sum_l f_{lk} \frac{ \int_{{\rm IV}_{il}} W (\vec \Omega ) \cdot d\vec \Omega }{ \int_{{\rm IV}}  W (\vec \Omega ) \cdot d\vec \Omega } \label{eq:discretize}\\
&\equiv & \sum_l  W_{il} f_{lk} \label{eq:Wf}
\end{eqnarray}
%%%
where $l$ is the index for longitudinal slices, $f_{lk}$ is the area fraction of the $k$-th surface type in the $l$-th longitudinal slice. 
The approximation is valid only when $f_k(\vec \Omega)$ does not change or changes little in the $l$-th slice for all $k$ and $l$. 
In the last equation, $W_{il}$ is the weight function for $i$-th epoch and $l$-th longitudinal slice which depends only on the observational geometry. 
As a result,
%%%
\begin{equation}
d_{ij} = R_p^2 \sum _{l,k} W_{il} \, f_{lk} \, s_{kj} \label{eq:d_f_s}
\end{equation}
%%%
with the following condition similar to equation (\ref{eq:cond_f_ast}):
%%%
\begin{eqnarray}
\begin{cases}
\;\; 0 < s_{kj} < 1 \;\;\; & \mbox{for any $k$, $j$} \\
\;\; 0 < f_{lk} < 1 \;\;\; & \mbox{for any $l$, $k$} \label{eq:cond_f} \\
\;\; \sum_k f_{lk} = 1 & \mbox{for any $l$} 
\end{cases}
\label{eq:constraints}
\end{eqnarray}
%%%
Now, the relevant problem is, given $d$, estimate $\{f, s\}$ subject to the constraints of equation (\ref{eq:constraints})---this is where \citet{Cowan2013} stood.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fit data with $f$ or with $f^{\ast}$?}
\label{ss:degeneracy}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We argue that estimating $\{ f, s \}$ is no superior to estimating $\{ f^{\ast }, s \}$. 
%
One reason comes from the approximation adopted in equation (\ref{eq:discretize}) is valid only when $f_k (\vec \Omega)$ does not change significantly in the $l$-th slice. 
Otherwise, fitting data with $f$ by equation (\ref{eq:Wf}) produces some errors by design.

%Suppose the above approximation is justified. 
In essence, $f^{\ast }_{ik}$ and $f_{lk}$ contains the same information, where they are related through a fixed $W_{il}$ (eq. \ref{eq:Wf}). 
%
If $L=I$, in general there is a one-to-one relationship between $f^{\ast}_{ik}$ and $f_{lk}$ through $W_{il}$ as far as ${\rm rank}(W)=L(=I)$. 
In this case, clearly, equations (\ref{eq:cond_f_ast}) and (\ref{eq:cond_f}) are equivalent, given $W_{il}>0$ (for any $i$ and $j$) and $\sum _l W_{il} = 1$. 
%
If $L<I$, the matrix $W$ works as a low-pass filter \citep{Cowan2013}. 
In this case, the inverse problem described by equation (\ref{eq:Wf}), $f^{\ast }=Wf$, is an overdetermined system, and one must to find the solution for $f$ by some means (e.g. the least squares method), and the resultant $f$ may not automatically satisfy the condition listed by eqution (\ref{eq:cond_f}). 
However, this is because of the design of the problem and the assumed lowered resolution.

@@

@@

As seen above, the solution is highly degenerate, and in general we cannot reduce to a unique solution. 
Therefore, we may want to give a prior in estimating the area fractions. 
What kind of assumptions can we make?

In general, the large time/longitude variations in apparent area fractions corresponds to spectrally similar surface types, because such solution correspond to the proximity of the inner boundary in the $J$-dimensional space discussed above. 
On the contrary, the spectrally separated surface types correspond to the small time/longitude variations in the apparent area fractions. 
In the real world, we might expect that the latter is more likely, i.e., covering area fraction of any surface type would not change significantly along the longitude (or in time). 
This would be a reasonable assumption for the estimation of $\{s, \fast \}$, since the illuminated and visible area is usually broad and tend to change the averaged area fraction of each surface type only gently. 
Although there is no physical proof for this assumption...

The assumption of the smoothness and the small variation of covering area fraction may be imposed using the concept of Gaussian Process \citep{Rasmussen2005}.

Essentially, the probability of the

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Bayesian Formulation}
\label{ss:regularization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%
\begin{eqnarray}
Posterior = \prod\limits_{i,j} \frac{1}{\sqrt{2\pi }\sigma_{ij}} &&  \exp \left[ - \frac{(d_{ij} - \sum _{lk} W_{il} f_{lk} s_{kj})^2}{2 \sigma _{ij}^2} \right] \notag \\
&& \times \; \mathcal{P} (\{f_{l,k}\}) \cdot \mathcal{P}(\{s_{l,k}\}) 
\end{eqnarray}
%%%
where $\mathcal{P} (X) $ is a prior for $X$.

In order to take account of constraints (equation (\ref{eq:constraints})), we adopt the following transformation of variables: 
%%%
\begin{eqnarray}
\begin{cases}
t_{kj} = \displaystyle \ln \left[ \frac{ s_{kj} }{1 - s_{kj} } \right]  \;\;\; & \mbox{for any $k$, $j$} \\
g_{lk} = \displaystyle \ln \left[ \frac{ f_{lk} }{1 - \sum _{m=0}^{k} f_{lm} } \right] \;\;\; & \mbox{for $k=1, ..., K-1$, any $j$} 
\end{cases}
\end{eqnarray}
%%%
Note that there is a 1-to-1 correspondence between $t$ and $s$ or $g$ and $f$ where $s$ and $f$ always satisfy our constraints.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!hbt]
    \begin{minipage}{0.5\hsize}
    \begin{center}
\includegraphics[width=\hsize]{IGBP_simplemap.pdf}
    \end{center}
     \end{minipage}
  \begin{minipage}{0.5\hsize}
    \begin{center}
    \includegraphics[width=\hsize]{mockdata_quadrature_bandsp.pdf}
    \end{center}
\end{minipage}
  \begin{minipage}{0.5\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_quadrature_factor.pdf}
    \end{center}
 \end{minipage}
   \begin{minipage}{0.5\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_quadrature_lc.pdf}
    \end{center}
 \end{minipage}
    \caption{Our mock data based on Earth surface. }
\label{fig:mockdata}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!bpth]
    \begin{center}
%\includegraphics[width=\hsize]{PCA_projected.pdf}
\includegraphics[width=\hsize]{mockdata_quadrature_PCplane.pdf}
    \end{center}
    \caption{Colors of 3 surface types and the light curves projected to the plane of 1st and 2nd principal components. }
\label{fig:shrinkwrap}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!bpth]
    \begin{center}
\includegraphics[width=\hsize]{xmed_std_time.pdf}
\includegraphics[width=\hsize]{xmed_std_longitude.pdf}
    \end{center}
    \caption{Left: result of estimation for $\{ \fast,\,s\}$. Right: result of estimation for $\{ f,\,s\}$}
\label{fig:shrinkwrap}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{rotating}
\usepackage{natbib}
\usepackage{enumerate}
\usepackage{multirow}
\usepackage{array}
\usepackage{appendix}
\usepackage{comment}
\usepackage{color,xcolor}
\usepackage{url}
\usepackage{here}
\usepackage{hyperref}
\hypersetup{colorlinks,linkcolor={blue!50!black},citecolor={blue!50!black},urlcolor={blue!50!black}}
\allowdisplaybreaks[1]
\bibliographystyle{apj}
\renewcommand{\bibname}{References}

@@

%%% http://www.oceanwave.jp/index.php?float%B4%C4%B6%AD(figure%2Ftable)%A4%CE%BD%D0%CE%CF%B0%CC%C3%D6%A4%F2%A5%B3%A5%F3%A5%C8%A5%ED%A1%BC%A5%EB

@@

By definition, the area fraction $\fast $ and reflectance spectra $s$ are subject to the following conditions:
%%%
\begin{eqnarray}
\begin{cases}
\;\; 0 < s_{kj} < 1 \;\;\; & \mbox{for any $k$, $j$} \\
\;\; 0 < \fast_{lk} < 1 \;\;\; & \mbox{for any $l$, $k$} \label{eq:cond_f_ast}\\
\;\; \sum_k \fast_{lk} = 1 & \mbox{for any $l$} 
\end{cases}
\end{eqnarray}
%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[tbh]
\caption{Indexes}
\begin{center}
\begin{tabular}{ccc} \hline \hline
Name & Symbol & Maximum \\ \hline
Observation Time & $i$ & I \\
Band & $j$ & J  \\
Surface Type & $k$ & K  \\
Longitudinal Slice  & $l$ & L \\ \hline
\end{tabular}
\end{center}
\label{tab:index}
\end{table}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The time variability of the apparent covering fraction $\fast $ due to spin rotation is related to the surface inhomogeneity along the equator. Approximately, $\fast _{ik}$ may be written as the weighted summation of the area fraction of $k$-th surface type in each of (the finite number $L$ of) longitudinal slices, i.e.,
%%%
\begin{eqnarray}
\fast _{ik} &=& \frac{ \int_{{\rm IV}_{i}} f_k (\vec \Omega ) \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }{ \int_{{\rm IV}_i}  \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }  \\
&=& \sum_l \frac{ \int_{{\rm IV}_{il}}  f_k (\vec \Omega ) W (\vec \Omega  ) \cdot d\vec \Omega }{ \int_{{\rm IV}}  W (\vec \Omega ) \cdot d\vec \Omega } \\
&\sim & \sum_l f_{lk} \frac{ \int_{{\rm IV}_{il}} W (\vec \Omega ) \cdot d\vec \Omega }{ \int_{{\rm IV}}  W (\vec \Omega ) \cdot d\vec \Omega } \label{eq:discretize}\\
&\equiv & \sum_l  W_{il} f_{lk} \label{eq:Wf}
\end{eqnarray}
%%%
where $l$ is the index for longitudinal slices, $f_{lk}$ is the area fraction of the $k$-th surface type in the $l$-th longitudinal slice. 
Strictly speaking, the approximation is valid only when $f_k(\vec \Omega)$ does not change or changes little across the $l$-th slice for all $k$ and $l$. 
In the last equation, $W_{il}$ is the weight function for $i$-th epoch and $l$-th longitudinal slice which depends only on the observational geometry. 
As a result,
%%%
\begin{equation}
d_{ij} = R_p^2 \sum _{l,k} W_{il} \, f_{lk} \, s_{kj} \label{eq:d_f_s}
\end{equation}
%%%
with the following condition similar to equation (\ref{eq:cond_f_ast}):
%%%
\begin{eqnarray}
\begin{cases}
\;\; 0 < s_{kj} < 1 \;\;\; & \mbox{for any $k$, $j$} \\
\;\; 0 < f_{lk} < 1 \;\;\; & \mbox{for any $l$, $k$} \label{eq:cond_f} \\
\;\; \sum_k f_{lk} = 1 & \mbox{for any $l$} 
\end{cases}
\label{eq:constraints}
\end{eqnarray}
%%%
Now, the relevant problem is, given $d$, estimate $\{f, s\}$ subject to the constraints of equation (\ref{eq:constraints})---this is where \citet{Cowan2013} stood.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fit data with $\fast $ or with $f$ ?}
\label{ss:degeneracy}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We argue that estimating $\{ f, s \}$ is no superior to estimating $\{ \fast, s \}$. 
%
One reason comes from the approximation adopted in equation (\ref{eq:discretize}) is valid only when $f_k (\vec \Omega)$ does not change significantly in the $l$-th slice. 
Otherwise, fitting data with $f$ by equation (\ref{eq:Wf}) produces some errors by design.

%Suppose the above approximation is justified. 
In essence, $f^{\ast }_{ik}$ and $f_{lk}$ contains the same information, where they are related through a fixed $W_{il}$ (eq. \ref{eq:Wf}). 
%
If $L=I$, in general there is a one-to-one relationship between $\fast_{ik}$ and $f_{lk}$ through $W_{il}$ as far as ${\rm rank}(W)=L(=I)$. 
In this case, clearly, equations (\ref{eq:cond_f_ast}) and (\ref{eq:cond_f}) are equivalent, given $W_{il}>0$ (for any $i$ and $j$) and $\sum _l W_{il} = 1$. 
%
If $L<I$, the matrix $W$ works as a low-pass filter \citep{Cowan2013}. 
In this case, the inverse problem described by equation (\ref{eq:Wf}), $\fast=Wf$, is an overdetermined system, and one must to find the solution for $f$ by some means (e.g. the least squares method), and the resultant $f$ may not automatically satisfy the condition listed by equation (\ref{eq:cond_f}).

However, this is because of the design of the problem and the assumed lowered resolution.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Bayesian Formulation ?}
\label{ss:regularization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Assuming the observational noise is distributed as Gaussian, 
%%%
\begin{eqnarray}
&&Posterior 
\\ &&= \prod\limits_{i,j}\left[  \frac{1}{\sqrt{2\pi }\sigma_{ij}} \exp \left\{ - \frac{(d_{ij} - \sum _{lk} W_{il} f_{lk} s_{kj})^2}{2 \sigma _{ij}^2} \right\} \right] \notag \\
&& \;\;\;\;\;\;\;\;\;\; \times \; \mathcal{P} (\{f_{lk}\}) \cdot \mathcal{P}(\{s_{kj}\}) 
\end{eqnarray}
%%%
where $\mathcal{P} (X) $ is a prior for $X$.

In order to take account of constraints (equation (\ref{eq:constraints})), we adopt the following transformation of variables: 
%%%
\begin{eqnarray}
\begin{cases}
t_{kj} = \displaystyle \ln \left[ \frac{ s_{kj} }{1 - s_{kj} } \right]  \;\;\; & \mbox{for any $k$, $j$} \\
g_{lk} = \displaystyle \ln \left[ \frac{ f_{lk} }{1 - \sum _{m=0}^{k} f_{lm} } \right] \;\;\; & \mbox{for $k=1, ..., K-1$} %, any $j$} 
\end{cases}
\end{eqnarray}
%%%
Note that there is a 1-to-1 correspondence between $t$ and $s$ or $g$ and $f$ where $s$.

@@

As seen in Section \ref{ss:degeneracy}, the solution is highly degenerate, and in general we cannot reduce to a unique solution. 
Therefore, we may want to give a prior in estimating the area fractions. 
What kind of assumptions can we make?

In general, the large time/longitude variations in apparent area fractions corresponds to spectrally similar surface types, because such solution correspond to the proximity of the inner boundary in the $J$-dimensional space discussed above. 
On the contrary, spectrally well-separated surface types correspond to small time/longitude variations in the apparent area fractions. 
In the real world, we might expect that the latter is more likely, i.e., covering area fraction of any surface type would not change significantly along the longitude (or in time). 
This would be a reasonable assumption for the estimation of $\{s, \fast \}$, since the illuminated and visible area is usually broad and tend to change the averaged area fraction of each surface type only gently. 
Although there is no physical proof for this assumption...

The assumption of the smoothness of the time variation of covering area fraction may be imposed using the concept of Gaussian Process \citep{Rasmussen2005}. 
We assume that the variation of the covering fraction follows Gaussian Process, which means the any finite number of the covering fractions have a joint Gaussian distribution. 
Such assumption can be represented by the prior of 
%%%
\begin{equation}
\mathcal{P}_{\fast} = \frac{1}{\sqrt{ 2 \pi }^{L(K-1)} \sqrt{| \tilde \Sigma |} } \exp( - \fast^{T} \tilde \Sigma ^{-1} \fast ) 
\end{equation}
%%%
or
%%%
\begin{equation}
\mathcal{P}_f = \frac{1}{\sqrt{ 2 \pi }^{L(K-1)} \sqrt{| \Sigma  |}} \exp( - f^{T} \Sigma ^{-1} f ) 
\end{equation}
%%%
where $\Sigma $ is the kernel representing the property of the gaussian processes assumed for $f$.
In this paper, we assume squared exponential covariance function, i.e., 
%%%
\begin{equation}
\tilde \Sigma _{ii'} = \exp \left( - \frac{|t_i- t_i'|^2}{2 (\Delta t_0) ^2} \right)
\end{equation}
%%%
where $t_i$ is the time of $i$-th observation and $(\Delta t_0)$ is characteristic time scale over which $\fast $ is supposedly correlated, or 
%%%
\begin{equation}
\Sigma _{ll'} = \exp \left( - \frac{|\phi_l- \phi_{l'}|^2}{2 \phi_0^2} \right)
\end{equation}
%%%
where $\phi_l$ is the longitude of $l$-th slice, and $\phi_0$ is the characteristic longitudes over which $f$ is correlated.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[hb!]
    \begin{minipage}{0.5\hsize}
    \begin{center}
\includegraphics[width=\hsize]{IGBP_simplemap.pdf}
    \end{center}
     \end{minipage}
  \begin{minipage}{0.5\hsize}
    \begin{center}
    \includegraphics[width=\hsize]{mockdata_quadrature_bandsp.pdf}
    \end{center}
\end{minipage}
  \begin{minipage}{0.5\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_quadrature_factor.pdf}
    \end{center}
 \end{minipage}
   \begin{minipage}{0.5\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_quadrature_lc.pdf}
    \end{center}
 \end{minipage}
    \caption{Our mock data based on Earth surface. }
\label{fig:mockdata}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\clearpage

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!bpth]
    \begin{center}
%\includegraphics[width=\hsize]{PCA_projected.pdf}
\includegraphics[width=\hsize]{mockdata_quadrature_PCplane_reduced.pdf}
    \end{center}
    \caption{Colors of 3 surface types and the light curves projected to the plane of 1st and 2nd principal components. }
\label{fig:shrinkwrap}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!htbp]
    \begin{center}
\includegraphics[width=\hsize]{xmed_std_time.pdf}
\includegraphics[width=\hsize]{xmed_std_longitude.pdf}
    \end{center}
    \caption{Left: result of estimation for $\{ \fast,\,s\}$. Right: result of estimation for $\{ f,\,s\}$}
\label{fig:shrinkwrap}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage
